### 2023-03-23 Paper/chapters/appendix.tex

\chapter{Appendix}

\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.h}]
#ifndef GIFT_VEC_SLICED_H
#define GIFT_VEC_SLICED_H

#include <stdint.h>
#include <arm_neon.h>

#define ROUNDS_GIFT_64 28

// expose for benchmarking
uint8x16_t shl(uint8x16_t v, int n);
uint8x16_t shr(uint8x16_t v, int n);
void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b,
                                 uint8x16_t m, int n);
void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2]);

void gift_64_vec_sliced_subcells(uint8x16x4_t cipher_state[restrict 2]);
void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cipher_state[restrict 2]);
void gift_64_vec_sliced_permute(uint8x16x4_t cipher_state[restrict 2]);
void gift_64_vec_sliced_permute_inv(uint8x16x4_t cipher_state[2]);
void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
                                            const uint64_t key[restrict 2]);

void gift_64_vec_sliced_init(void);

void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
                                const uint64_t m[restrict 16],
                                const uint64_t key[restrict 2]);
void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
                                const uint64_t c[restrict 16],
                                const uint64_t key[restrict 2]);

#endif
\end{lstlisting}

\begin{lstlisting}[language=c, caption={gift\_vec\_sliced.c}]
#include "gift_vec_sliced.h"

#include <stddef.h>
#include <stdio.h>
#include <string.h>

static uint64_t perm_u64[16] = {
        0x1410050115110400UL, 0x1410050115110400UL,
        0x0501151104001410UL, 0x0501151104001410UL,
        0x1511040014100501UL, 0x1511040014100501UL,
        0x0400141005011511UL, 0x0400141005011511UL,
        0x1612070317130602UL, 0x1612070317130602UL,
        0x0703171306021612UL, 0x0703171306021612UL,
        0x1713060216120703UL, 0x1713060216120703UL,
        0x0602161207031713UL, 0x0602161207031713UL
};

static uint64_t perm_inv_u64[16] = {
        0x1511050114100400UL, 0x1511050114100400UL,
        0x1713070316120602UL, 0x1713070316120602UL,
        0x1115010510140004UL, 0x1115010510140004UL,
        0x1317030712160206UL, 0x1317030712160206UL,
        0x1317030712160206UL, 0x1317030712160206UL,
        0x1511050114100400UL, 0x1511050114100400UL,
        0x1713070316120602UL, 0x1713070316120602UL,
        0x1115010510140004UL, 0x1115010510140004UL
};

static uint8x16x4_t perm[2];
static uint8x16x4_t perm_inv[2];

static uint8x16_t pack_mask_0;
static uint8x16_t pack_mask_1;
static uint8x16_t pack_mask_2;

static const int round_constant[] = {
        // rounds 0-15
        0x01, 0x03, 0x07, 0x0F, 0x1F, 0x3E, 0x3D, 0x3B,
        0x37, 0x2F, 0x1E, 0x3C, 0x39, 0x33, 0x27, 0x0E,
        // rounds 16-31
        0x1D, 0x3A, 0x35, 0x2B, 0x16, 0x2C, 0x18, 0x30,
        0x21, 0x02, 0x05, 0x0B, 0x17, 0x2E, 0x1C, 0x38,
        // rounds 32-47
        0x31, 0x23, 0x06, 0x0D, 0x1B, 0x36, 0x2D, 0x1A,
        0x34, 0x29, 0x12, 0x24, 0x08, 0x11, 0x22, 0x04
};

uint8x16_t shl(uint8x16_t v, int n)
{
        uint64_t l[2];
        vst1q_u64(l, v);
        l[1] = l[1] << n | (l[0] >> (64 - n));
        l[1] <<= n;
        return vreinterpretq_u8_u64(vld1q_u64(l));
}

uint8x16_t shr(uint8x16_t v, int n)
{
        uint64_t l[2];
        vst1q_u64(l, v);
        l[0] = l[0] >> n | (((l[1] << (64 - n)) >> (64 - n)) << (64 - n));
        l[1] >>= n;
        return vreinterpretq_u8_u64(vld1q_u64(l));
}

void gift_64_vec_sliced_swapmove(uint8x16_t *restrict a, uint8x16_t *restrict b, uint8x16_t m, int n)
{

        uint8x16_t t = vandq_u8(veorq_u8(shr(*a, n), *b), m);
        *b = veorq_u8(*b, t);
        *a = veorq_u8(*a, shl(t, n));
}

void gift_64_vec_sliced_bits_pack(uint8x16x4_t m[restrict 2])
{
        // take care not to shift mask bits out of the register
        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[1], pack_mask_0, 1);
        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[0].val[3], pack_mask_0, 1);
        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[1], pack_mask_0, 1);
        gift_64_vec_sliced_swapmove(&m[1].val[2], &m[1].val[3], pack_mask_0, 1);

        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[0].val[2], pack_mask_1, 2);
        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[0].val[3], pack_mask_1, 2);
        gift_64_vec_sliced_swapmove(&m[1].val[0], &m[1].val[2], pack_mask_1, 2);
        gift_64_vec_sliced_swapmove(&m[1].val[1], &m[1].val[3], pack_mask_1, 2);

        // make bytes (a0 b0 c0 d0 a4 b4 c4 d4 -> a0 b0 c0 d0 e0 f0 g0 h0)
        gift_64_vec_sliced_swapmove(&m[0].val[0], &m[1].val[0], pack_mask_2, 4);
        gift_64_vec_sliced_swapmove(&m[0].val[2], &m[1].val[2], pack_mask_2, 4);
        gift_64_vec_sliced_swapmove(&m[0].val[1], &m[1].val[1], pack_mask_2, 4);
        gift_64_vec_sliced_swapmove(&m[0].val[3], &m[1].val[3], pack_mask_2, 4);
}

void gift_64_vec_sliced_subcells(uint8x16x4_t cs[restrict 2])
{
        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));
        uint8x16_t t = veorq_u8(cs[0].val[0], vandq_u8(cs[0].val[1], cs[0].val[3]));
        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
        cs[0].val[0] = veorq_u8(cs[0].val[3], cs[0].val[2]);
        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
        cs[0].val[3] = t;

        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
        t            = veorq_u8(cs[1].val[0], vandq_u8(cs[1].val[1], cs[1].val[3]));
        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
        cs[1].val[0] = veorq_u8(cs[1].val[3], cs[1].val[2]);
        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
        cs[1].val[3] = t;
}

void gift_64_vec_sliced_subcells_inv(uint8x16x4_t cs[restrict 2])
{
        uint8x16_t t = cs[0].val[3];
        cs[0].val[2] = veorq_u8(cs[0].val[2], vandq_u8(t, cs[0].val[1]));
        cs[0].val[0] = vmvnq_u8(cs[0].val[0]);
        cs[0].val[1] = veorq_u8(cs[0].val[1], cs[0].val[0]);
        cs[0].val[3] = veorq_u8(cs[0].val[0], cs[0].val[2]);
        cs[0].val[2] = veorq_u8(cs[0].val[2], vorrq_u8(t, cs[0].val[1]));
        cs[0].val[0] = veorq_u8(t, vandq_u8(cs[0].val[1], cs[0].val[3]));
        cs[0].val[1] = veorq_u8(cs[0].val[1], vandq_u8(cs[0].val[0], cs[0].val[2]));

        t            = cs[1].val[3];
        cs[1].val[2] = veorq_u8(cs[1].val[2], vandq_u8(t, cs[1].val[1]));
        cs[1].val[0] = vmvnq_u8(cs[1].val[0]);
        cs[1].val[1] = veorq_u8(cs[1].val[1], cs[1].val[0]);
        cs[1].val[3] = veorq_u8(cs[1].val[0], cs[1].val[2]);
        cs[1].val[2] = veorq_u8(cs[1].val[2], vorrq_u8(t, cs[1].val[1]));
        cs[1].val[0] = veorq_u8(t, vandq_u8(cs[1].val[1], cs[1].val[3]));
        cs[1].val[1] = veorq_u8(cs[1].val[1], vandq_u8(cs[1].val[0], cs[1].val[2]));
}

void gift_64_vec_sliced_permute(uint8x16x4_t cs[restrict 2])
{
        uint8x16x2_t pairs[4] = {
                { .val = { cs[0].val[0], cs[1].val[0] }},
                { .val = { cs[0].val[1], cs[1].val[1] }},
                { .val = { cs[0].val[2], cs[1].val[2] }},
                { .val = { cs[0].val[3], cs[1].val[3] }},
        };

        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm[0].val[0]);
        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm[0].val[1]);
        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm[0].val[2]);
        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm[0].val[3]);

        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm[1].val[0]);
        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm[1].val[1]);
        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm[1].val[2]);
        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm[1].val[3]);
}

void gift_64_vec_sliced_permute_inv(uint8x16x4_t cs[restrict 2])
{
        uint8x16x2_t pairs[4] = {
                { .val = { cs[0].val[0], cs[1].val[0] }},
                { .val = { cs[0].val[1], cs[1].val[1] }},
                { .val = { cs[0].val[2], cs[1].val[2] }},
                { .val = { cs[0].val[3], cs[1].val[3] }},
        };

        cs[0].val[0] = vqtbl2q_u8(pairs[0], perm_inv[0].val[0]);
        cs[0].val[1] = vqtbl2q_u8(pairs[1], perm_inv[0].val[1]);
        cs[0].val[2] = vqtbl2q_u8(pairs[2], perm_inv[0].val[2]);
        cs[0].val[3] = vqtbl2q_u8(pairs[3], perm_inv[0].val[3]);

        cs[1].val[0] = vqtbl2q_u8(pairs[0], perm_inv[1].val[0]);
        cs[1].val[1] = vqtbl2q_u8(pairs[1], perm_inv[1].val[1]);
        cs[1].val[2] = vqtbl2q_u8(pairs[2], perm_inv[1].val[2]);
        cs[1].val[3] = vqtbl2q_u8(pairs[3], perm_inv[1].val[3]);
}

void gift_64_vec_sliced_generate_round_keys(uint8x16x4_t round_keys[restrict ROUNDS_GIFT_64][2],
                                            const uint64_t key[restrict 2])
{
        // TODO shift-and-or

        uint64_t key_state[] = {key[0], key[1]};
        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
                int v = (key_state[0] >> 0 ) & 0xffff;
                int u = (key_state[0] >> 16) & 0xffff;

                // add round key (RK=U||V)
                uint64_t round_key[8] = { 0x0UL };
                for (size_t i = 0; i < 8; i++) {
                        int key_bit_v   = (v >> (2 * i + 0)) & 0x1;
                        int key_bit_u   = (u >> (2 * i + 0)) & 0x1;
                        round_key[0] ^= (uint64_t)key_bit_v << (i * 8);
                        round_key[1] ^= (uint64_t)key_bit_u << (i * 8);

                        key_bit_v   = (v >> (2 * i + 1)) & 0x1;
                        key_bit_u   = (u >> (2 * i + 1)) & 0x1;
                        round_key[4] ^= (uint64_t)key_bit_v << (i * 8);
                        round_key[5] ^= (uint64_t)key_bit_u << (i * 8);
                }

                // add single bit
                round_key[7] ^= 1UL << (7 * 8);

                // add round constants
                round_key[3] ^= ((round_constant[round] >> 0) & 0x1) << (0 * 8);
                round_key[7] ^= ((round_constant[round] >> 1) & 0x1) << (0 * 8);
                round_key[3] ^= ((round_constant[round] >> 2) & 0x1) << (1 * 8);
                round_key[7] ^= ((round_constant[round] >> 3) & 0x1) << (1 * 8);
                round_key[3] ^= ((round_constant[round] >> 4) & 0x1) << (2 * 8);
                round_key[7] ^= ((round_constant[round] >> 5) & 0x1) << (2 * 8);

                // extend bits to bytes
                for (size_t i = 0; i < 8; i++) {
                        round_key[i] |= round_key[i] << 1;
                        round_key[i] |= round_key[i] << 2;
                        round_key[i] |= round_key[i] << 4;
                }

                // load into vector registers (upper and lower part are same)
                round_keys[round][0].val[0] = vdupq_n_u64(round_key[0]);
                round_keys[round][0].val[1] = vdupq_n_u64(round_key[1]);
                round_keys[round][0].val[2] = vdupq_n_u64(round_key[2]);
                round_keys[round][0].val[3] = vdupq_n_u64(round_key[3]);
                round_keys[round][1].val[0] = vdupq_n_u64(round_key[4]);
                round_keys[round][1].val[1] = vdupq_n_u64(round_key[5]);
                round_keys[round][1].val[2] = vdupq_n_u64(round_key[6]);
                round_keys[round][1].val[3] = vdupq_n_u64(round_key[7]);

                // update key state
                int k0 = (key_state[0] >> 0 ) & 0xffffUL;
                int k1 = (key_state[0] >> 16) & 0xffffUL;
                k0 = (k0 >> 12) | ((k0 & 0xfff) << 4);
                k1 = (k1 >> 2 ) | ((k1 & 0x3  ) << 14);
                key_state[0] >>= 32;
                key_state[0] |= (key_state[1] & 0xffffffffUL) << 32;
                key_state[1] >>= 32;
                key_state[1] |= ((uint64_t)k0 << 32) | ((uint64_t)k1 << 48);
        }
}

void gift_64_vec_sliced_init(void)
{
        // permutations
        perm[0] = vld1q_u8_x4((uint8_t*)&perm_u64[0]);
        perm[1] = vld1q_u8_x4((uint8_t*)&perm_u64[8]);

        // inverse permutations
        perm_inv[0] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[0]);
        perm_inv[1] = vld1q_u8_x4((uint8_t*)&perm_inv_u64[8]);

        // packing masks
        pack_mask_0 = vdupq_n_u8(0x55);
        pack_mask_1 = vdupq_n_u8(0x33);
        pack_mask_2 = vdupq_n_u8(0x0f);
}

void gift_64_vec_sliced_encrypt(uint64_t c[restrict 16],
                                const uint64_t m[restrict 16],
                                const uint64_t key[restrict 2])
{
        uint8x16x4_t s[2];
        s[0] = vld1q_u8_x4((uint8_t*)&m[0]);
        s[1] = vld1q_u8_x4((uint8_t*)&m[8]);
        gift_64_vec_sliced_bits_pack(s);

        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
        gift_64_vec_sliced_generate_round_keys(round_keys, key);

        for (int round = 0; round < ROUNDS_GIFT_64; round++) {
                gift_64_vec_sliced_subcells(s);
                gift_64_vec_sliced_permute(s);

                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);
        }

        gift_64_vec_sliced_bits_pack(s);
        vst1q_u8_x4((uint8_t*)&c[0], s[0]);
        vst1q_u8_x4((uint8_t*)&c[8], s[1]);
}

void gift_64_vec_sliced_decrypt(uint64_t m[restrict 16],
                                const uint64_t c[restrict 16],
                                const uint64_t key[restrict 2])
{
        uint8x16x4_t s[2];
        s[0] = vld1q_u8_x4((uint8_t*)&c[0]);
        s[1] = vld1q_u8_x4((uint8_t*)&c[8]);
        gift_64_vec_sliced_bits_pack(s);

        uint8x16x4_t round_keys[ROUNDS_GIFT_64][2];
        gift_64_vec_sliced_generate_round_keys(round_keys, key);

        for (int round = ROUNDS_GIFT_64 - 1; round >= 0; round--) {
                s[0].val[0] = veorq_u8(s[0].val[0], round_keys[round][0].val[0]);
                s[0].val[1] = veorq_u8(s[0].val[1], round_keys[round][0].val[1]);
                s[0].val[2] = veorq_u8(s[0].val[2], round_keys[round][0].val[2]);
                s[0].val[3] = veorq_u8(s[0].val[3], round_keys[round][0].val[3]);
                s[1].val[0] = veorq_u8(s[1].val[0], round_keys[round][1].val[0]);
                s[1].val[1] = veorq_u8(s[1].val[1], round_keys[round][1].val[1]);
                s[1].val[2] = veorq_u8(s[1].val[2], round_keys[round][1].val[2]);
                s[1].val[3] = veorq_u8(s[1].val[3], round_keys[round][1].val[3]);

                gift_64_vec_sliced_permute_inv(s);
                gift_64_vec_sliced_subcells_inv(s);
        }

        gift_64_vec_sliced_bits_pack(s);
        vst1q_u8_x4((uint8_t*)&m[0], s[0]);
        vst1q_u8_x4((uint8_t*)&m[8], s[1]);
}
\end{lstlisting}