### 2023-03-23 Paper/paper.tex

\documentclass[12pt]{report}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{ulem} % underline
\usepackage{float} % H float specifier
\usepackage{hyperref} % hyperlinks
\usepackage{tabularx} % tabular with column specifier X to break lines
% \usepackage{hyphenat}
\usepackage{bookmark}
\usepackage{systeme} % systems of equation
\usepackage[utf8]{inputenc} % necessary for proper unicode support
\usepackage{parcolumns} % multiple column parallel typesetting
\usepackage[english]{babel} % comment this for Angelsächsisch
\usepackage{graphicx} % including graphics
\usepackage{sourcecodepro} % best font
\usepackage{listings} % source code listings
\usepackage{biblatex} % better citation
% \usepackage[a4paper, top=1cm, left=2cm, right=2cm, includeheadfoot]{geometry} % customize page geometry

% umlauts in lstlistings env
\lstset{
    literate={ö}{{\"o}}1
    {ä}{{\"a}}1
    {ü}{{\"u}}1
    {Ö}{{\"O}}1
    {Ä}{{\"A}}1
    {Ü}{{\"U}}1
}

\lstset{
    basicstyle=\scriptsize\ttfamily,
    breaklines=true
}

\lstset {
    frame=single,
    numbers=left,
    stepnumber=1,
    firstnumber=1,
    numberfirstline=true,
    texcl=false
}

\providecommand{\lxor}{\veebar}
\renewcommand{\proofname}{Beweis.}
\newcommand{\sref}[1]{\textsuperscript{\ref{#1}}}
\newcommand{\F}[1]{\mathbb{F}_2^{#1}}

\addbibresource{refs.bib}

\title{
    \textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\
    {\footnotesize Bachelorarbeit}
}
\author{Bastian Engel}
\date{\today}

\begin{document}

\maketitle

\include{chapters/abstract}
\include{chapters/declaration}

\tableofcontents

\include{chapters/introduction}
\include{chapters/implementation-strategies}
\include{chapters/implementation}
\include{chapters/evaluation}
\include{chapters/acknowledgements}
\include{chapters/appendix}

\printbibliography

\end{document}