### 2023-04-04 Paper/chapters/implementation.tex

\chapter{Implementation}

Implementations in the C programming language for the presented strategies can
be found in Appendix \ref{app:cimpl}. Although directly writing Assembler code could
result in a small performance benefit, this generally increases the work
necessary by an order of magnitude for only limited results. Instruction-level
optimization and in particular register allocation is left to the compiler.