### 2023-05-18 Paper/chapters/appendix.tex
rename files, add implementation sample code

\appendix
\chapter{C implementations}
\label{app:cimpl}

\section{Implementations for SPN}
\subsection{Table-based}

\lstinputlisting[language=c, caption={\texttt{gift\_table.h}}]{../impl/gift/table.h}
\lstinputlisting[language=c, caption={\texttt{gift\_table.c}}]{../impl/gift/table.c}

\subsection{Using \texttt{vperm}}

\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.h}}]{../impl/gift/vec_sbox.h}
\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sbox.c}}]{../impl/gift/vec_sbox.c}

\subsection{Bitslicing}

\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.h}}]{../impl/gift/vec_sliced.h}
\lstinputlisting[language=c, caption={\texttt{gift\_vec\_sliced.c}}]{../impl/gift/vec_sliced.c}

\chapter{Lorem dolor}