### 2023-05-19 Paper/chapters/appendix.tex
finish draft

\appendix
\chapter{Detailed benchmarking results}
\label{app:benchmarks}

\section{GIFT}

\begin{table}[!htbp]
    \centering
    \caption{Benchmarks for GIFT}
    \scriptsize
    \begin{tabular}{llcccc}
        \toprule
        & & \multicolumn{2}{c}{Cortex-A53} & \multicolumn{2}{c}{Cortex-A73} \\
        \cmidrule(lr){3-4}\cmidrule(lr){5-6}
        Strategy & Component & $lat$ (c/B) & $thr$ (MiB/s) & $lat$ (c/B) & $thr$ (MiB/s) \\
        \midrule
        \multirow{2}{*}{Naive GIFT-64} & \texttt{round\_keys} & 223.54 & \multirow{2}{*}{1.22} & 190.34 & \multirow{2}{*}{2.33} \\
                                                & \texttt{encrypt} & 1367.66 & &  830.49 & \\
        \cmidrule(lr){2-2}
                                                & \texttt{ subcells} & 4.64 & & 3.13 & \\
                                                & \texttt{  permute} & 36.68 & & 21.09 & \\
        \midrule
        \multirow{2}{*}{Naive GIFT-128} & \texttt{round\_keys} & 182.91 & \multirow{2}{*}{0.51} & 167.16 & \multirow{2}{*}{0.79} \\
                                                 & \texttt{encrypt} & 3532.52 & & 2615.89 & \\
        \cmidrule(lr){2-2}
                                                 & \texttt{ subcells} & 6.91 & & 2.73 & \\
                                                 & \texttt{ permute} & 82.68 & & 61.27 & \\
        \midrule
        \multirow{2}{*}{Table-driven} & \texttt{round\_keys} & 223.81 & \multirow{2}{*}{13.59} & 190.11 & \multirow{2}{*}{16.09} \\
                                      & \texttt{encrypt} & 122.11 & & 119.62 & \\
        \cmidrule(lr){2-2}
                                      & \texttt{ subperm} & 5.15 & & 4.47 & \\
        \midrule
        \multirow{2}{*}{\texttt{vperm} S-box} & \texttt{round\_keys} & 308.12 & \multirow{2}{*}{1.10} & 270.69 & \multirow{2}{*}{1.58} \\
                                              & \texttt{encrypt} & 1514.28 & & 1218.18 & \\
        \cmidrule(lr){2-2}
                                              & \texttt{ subcells} & 1.63 & & 1.13 & \\
                                              & \texttt{ permute} & 53.18 & & 42.40 & \\
                                              & \texttt{ pack} & 7.51 & & 1.24 & \\
                                              & \texttt{ unpack} & 7.39 & & 4.67 & \\
        \midrule
        \multirow{2}{*}{Bitsliced} & \texttt{round\_keys} & 19.93 & \multirow{2}{*}{108.80} & 16.12 & \multirow{2}{*}{150.50} \\
                                   & \texttt{encrypt} & 16.69 & & 13.98 & \\
        \cmidrule(lr){2-2}
                                   & \texttt{ subcells} & 0.41 & & 0.06 & \\
                                   & \texttt{ permute} & 0.39 & & 0.18 & \\
                                   & \texttt{ pack} & 1.81 & & 1.34 & \\
                                   & \texttt{ unpack} & 1.68 & & 1.31 & \\
        \bottomrule
    \end{tabular}
\end{table}

\section{Camellia}

\begin{table}[!htpb]
    \centering
    \caption{Benchmarks for Camellia}
    \scriptsize
    \begin{tabular}{llcccc}
        \toprule
        & & \multicolumn{2}{c}{Cortex-A53} & \multicolumn{2}{c}{Cortex-A73} \\
        \cmidrule(lr){3-4}\cmidrule(lr){5-6}
        Strategy & Component & $lat$ (c/B) & $thr$ (MiB/s) & $lat$ (c/B) & $thr$ (MiB/s) \\
        \midrule
        \multirow{2}{*}{Naive Camellia-128} & \texttt{round\_keys} & 15.26 & \multirow{2}{*}{33.74} & 11.08 & \multirow{2}{*}{46.71} \\
                                                & \texttt{encrypt} & 53.44 & &  43.51 & \\
        \cmidrule(lr){2-2}
                                                & \texttt{ feistel\_round} & 4.02 & & 2.66 & \\
                                                & \texttt{ S} & 2.01 & & 0.94 & \\
                                                & \texttt{ F} & 3.13 & & 2.25 & \\
                                                & \texttt{ P} & 2.03 & & 1.63 & \\
                                                & \texttt{ FL} & 1.06 & & 0.72 & \\
        \midrule
        \multirow{2}{*}{Naive Camellia-256} & \texttt{round\_keys} & 22.01 & \multirow{2}{*}{25.50} & 16.99 & \multirow{2}{*}{35.16} \\
                                                 & \texttt{encrypt} & 70.55 & & 58.27 & \\
        \midrule
        \multirow{2}{*}{Optimized Camellia-128} & \texttt{round\_keys} & 8.18 & \multirow{2}{*}{77.07} & 5.93 & \multirow{2}{*}{107.31} \\
                                                & \texttt{encrypt} & 23.63 & & 19.83 & \\
        \cmidrule(lr){2-2}
                                                & \texttt{ feistel\_round} & 2.32 & & 1.56 & \\
                                                & \texttt{ F} & 2.21 & & 1.43 & \\
                                                & \texttt{ FL} & 1.06 & & 0.69 & \\
        \midrule
        \multirow{2}{*}{Bytesliced Camellia-128} & \texttt{round\_keys} & 2.59 & \multirow{2}{*}{106.20} & 2.19 & \multirow{2}{*}{138.42} \\
                                                & \texttt{encrypt} & 17.14 & & 15.19 & \\
        \cmidrule(lr){2-2}
                                                & \texttt{ feistel\_round} & 1.00 & & 0.69 & \\
                                                & \texttt{ F} & 0.79 & & 0.58 & \\
                                                & \texttt{ FL} & 0.36 & & 0.12 & \\
                                                & \texttt{ pack} & 1.04 & & 0.96 & \\
                                                & \texttt{ unpack} & 0.91 & & 0.83 & \\
        \bottomrule
    \end{tabular}
\end{table}

\chapter{C source code}
\label{app:cimpl}

\section{GIFT}
\subsection{Table-based}

\lstinputlisting[language=c, caption={\texttt{gift/table.h}}]{../impl/gift/table.h}
\lstinputlisting[language=c, caption={\texttt{gift/table.c}}]{../impl/gift/table.c}

\subsection{Using \texttt{vperm}}

\lstinputlisting[language=c, caption={\texttt{gift/vec\_sbox.h}}]{../impl/gift/vec_sbox.h}
\lstinputlisting[language=c, caption={\texttt{gift/vec\_sbox.c}}]{../impl/gift/vec_sbox.c}

\subsection{Bitslicing}

\lstinputlisting[language=c, caption={\texttt{gift/vec\_sliced.h}}]{../impl/gift/vec_sliced.h}
\lstinputlisting[language=c, caption={\texttt{gift/vec\_sliced.c}}]{../impl/gift/vec_sliced.c}

\section{Camellia}

\subsection{Optimized non-SIMD}

\lstinputlisting[language=c, caption={\texttt{camellia/spec\_opt.h}}]{../impl/camellia/spec_opt.h}
\lstinputlisting[language=c, caption={\texttt{camellia/spec\_opt.c}}]{../impl/camellia/spec_opt.c}

\subsection{Byteslicing}

\lstinputlisting[language=c, caption={\texttt{camellia/bytesliced.h}}]{../impl/camellia/bytesliced.h}
\lstinputlisting[language=c, caption={\texttt{camellia/bytesliced.c}}]{../impl/camellia/bytesliced.c}