### 2023-05-20 Paper/chapters/abstract.tex
add abstract

\chapter*{Abstract}

Processor designs licensed by ARM are ubiquitous in many areas due to their low
power consumption and heat generation. Being used in nearly all modern
smartphones and tablets as well as in many desktop computers and IoT devices,
the need for securing communications and confidentiality through encryption
arises. Designing cryptographic algorithms for such low-power environments is
an active area of research and various designs have been proposed. We examine
in detail the two block ciphers GIFT and Camellia. For that, we first give an
introduction of the designs as well as the ARMv8 platform we develop for. We
then present multiple implementation strategies which leverage ARMv8 SIMD
processing facilities and realize these in the C programming language using
NEON intrinsics. In the end, we present performance benchmarks and a brief
comparison to conclude this work. We find bit- and bytesliced implementations
employing vector instructions to be the fastest.