### 2023-05-30 Paper/chapters/implementation-strategies.tex
finalize

\chapter{Implementation strategies}

Due to the structural differences of SPN- and Feistel network-based ciphers, we
shall analyze these two separately.

\section{Strategies for GIFT}

Three implementation strategies for substitution-permutation networks are
introduced by \cite{implx86:2014}:

\begin{itemize}
    \item Table-based implementations
    \item \texttt{vperm} implementations
    \item Bitslice implementations
\end{itemize}

\subsection{Table-based}

Table-driven programming is a simple way to increase performance of operations
by tabulating the results, therefore requiring only a single memory access to
acquire the result. This approach is obviously limited to manageable table
sizes, so while tabulating a function like the AES S-box
$S_{AES}:\F{8}\rightarrow \F{8}$ requires only $2^{11}$ space, tabulating the
GIFT permutation layer $P_{GIFT}:\F{64}\rightarrow \F{64}$ would require
$2^{70}$ space, which is totally unfeasible.

A common approach is to tabulate the output of each S-box, including the
diffusion layer, and then XORing the results together. Let $n$ denote the
internal cipher state size and $s$ the size of a single S-box in bits. For each
S-box $S_i,i\in\{0,\dots,\frac{n}{s}\}$, we can construct a mapping
$T_i:\F{s}\rightarrow \F{n}$ representing substitution with subsequent
permutation of that single S-box. The cipher state before round key addition is
then given by $\bigoplus_{i=0}^{\frac{n}{s}-1}{T_i(m_i)}$ for each $s$-bit
message chunk $m_i$. This approach requires space of
$\frac{n}{s}|\F{s}|n=\frac{n^2 2^s}{s}$ bits, which, for GIFT-64,
results in a manageable size of $\frac{64^2 2^4}{4}=2^{14}$ bits which equals
$16$ KiB.

\subsubsection{Constructing the tables}

For GIFT-64, table construction is relatively straightforward and can
be done as follows:

\begin{lstlisting}[caption={Table construction algorithm}, label={lst:tables}, escapechar=|]
    tables <- [][]
    for sbox_index from 0 to 15 do
        for sbox_input from 0 to 15 do|\label{lst:tablesbox}|
            output <- sbox(sbox_input)
            output <- permute(output << (4 * sbox_index))
            tables[sbox_index][sbox_input] <- output
\end{lstlisting}

Implementing this algorithm gives us the following table representing the first
and second S-box.

\[
    \begin{array}{l|l|l|c}
        x & T_0(x) & T_1(x) & \dots \\
        \hline
        0x0 & 0x1               & 0x1000000000000   & \dots \\
        0x1 & 0x8000000020000   & 0x800000002       & \dots \\
        0x2 & 0x400000000       & 0x40000           & \dots \\
        0x3 & 0x8000400000000   & 0x800040000       & \dots \\
        0x4 & 0x400020000       & 0x40002           & \dots \\
        0x5 & 0x8000400020001   & 0x1000800040002   & \dots \\
        0x6 & 0x20001           & 0x1000000000002   & \dots \\
        0x7 & 0x8000000000001   & 0x1000800000000   & \dots \\
        0x8 & 0x20000           & 0x2               & \dots \\
        0x9 & 0x8000400000001   & 0x1000800040000   & \dots \\
        0xa & 0x8000000020001   & 0x1000800000002   & \dots \\
        0xb & 0x400020001       & 0x1000000040002   & \dots \\
        0xc & 0x400000001       & 0x1000000040000   & \dots \\
        0xd & 0x0               & 0x0               & \dots \\
        0xe & 0x8000000000000   & 0x800000000       & \dots \\
        0xf & 0x8000400020000   & 0x800040002       & \dots
    \end{array}
\]

The tables for GIFT-128 can be generated in a similar way by looping
through all 32 S-boxes instead of 16 on line \ref{lst:tablesbox}.

\subsection{Using \texttt{vperm}}

The plenitude of different processing instructions introduced by
NEON\ref{ss:neon} allow flexible ways to further speed up algorithms having
reached their optimizational limit on non-SIMD platforms. \texttt{vperm}, a
general term standing for \textit{vector permute}, is a common instruction on
SIMD machines. Called \texttt{TBL} on NEON, it is used for parallel table
lookups and arbitrary permutations. It takes two inputs to perform a lanewise
lookup:

\begin{enumerate}
    \item A register with lookup values
    \item One or more registers containing data
\end{enumerate}

\subsubsection{S-box lookup}

This instruction can be used to implement S-box lookup of all 16 S-boxes in a
single instruction. We do this by packing our 64-bit cipher state
$s=s_{15}||s_{14}||\dots||s_0$ into a vector register $V_0$. Because we can
only operate on whole bytes, we put each 4-bit S-box into an 8-bit lane which
neatly fits into the 128-bit registers. We then put the S-box itself into
register $V_1$ which will be used as the data register for the table lookup.

The confusion layer can now be performed through one \texttt{TBL} instruction:

\begin{center}
    \texttt{TBL V0.16B, V1.16B, V0.16B}
\end{center}

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{Figures/tbl_example.pdf}
    \caption{Performing the S-Box lookup in parallel}
\end{figure}

\subsection{Bitslicing}

Bitsliced implementation techniques were first introduced to improve the
performance of DES in 1997 and work by viewing a processor with $n$-bit
registers as a machine capable of executing $n$ bitwise operations at
once \cite{desslicing:1997}. Bitslicing offers a performance advantage by
splitting up $n$ bits into $m$ slices to achieve a more efficient
representation which can exploit this bitwise parallelism. The structure of
GIFT naturally offers possibilities for bitslicing. We split the
cipher state bits $b_{63}b_{62}\dots b_0$ into four slices $S_i,
i\in\{0,1,2,3\}$ such that the $i$-th slice contains all $i$-th bits of the
individual S-boxes. This is equivalent to transposing the bit matrix.

\[
    S=\begin{bmatrix}
        S_0\\
        S_1\\
        S_2\\
        S_3
    \end{bmatrix}
    =\begin{bmatrix}
        b_{60}b_{56}b_{52}\dots b_0\\
        b_{61}b_{57}b_{53}\dots b_1\\
        b_{62}b_{58}b_{54}\dots b_2\\
        b_{63}b_{59}b_{55}\dots b_3
    \end{bmatrix}
\]

\subsubsection{Parallel S-Boxes}

This representation offers multiple advantages. We first note that computation
of the S-box can be executed in parallel, similar to the \texttt{vperm}
technique above. This can be done by finding a bitwise instruction sequence to
apply the S-box which has already been proposed by the original GIFT authors:

\begin{align*}
    S_1&\leftarrow S_1\oplus (S_0\land S_2) \\
    t&\leftarrow S_0\oplus (S_1\land S_3) \\
    S_2&\leftarrow S_2\oplus (t\lor S_1) \\
    S_0&\leftarrow S_3\oplus S_2 \\
    S_1&\leftarrow S_1\oplus S_0 \\
    S_0&\leftarrow \lnot S_0 \\
    S_2&\leftarrow S_2\oplus (t\land S_1) \\
    S_3&\leftarrow t
\end{align*}

This is very efficient as it only requires six XOR-, three AND and one OR
operation.

An important property of the permutation is the fact that bits always stay in
their slice. This means we can decompose the permutation $P$ into four
permutations $P_i,i\in\{0,1,2,3\}$ and apply these permutations
separately to each slice. One possible way to implement a permutation $P_i$ in
software is to mask off all bits individually, shift them to their correct
position and OR them together:

\[
    P_i(S_i)=\bigvee_{k=0}^{15}{(S_i\land m_i) \ll s_i}
\]

This approach requires $47$ operations, meaning all four permutations require
over $150$ operations which would present a major bottleneck to the round
function. We can improve on this by working on multiple message blocks at once
and using the aforementioned \texttt{vperm} instruction to implement the bit
shuffling. We then need only four instructions for the complete diffusion
layer.

\subsubsection{Using \texttt{vperm} for slice permutation}

We cannot use the \texttt{TBL} instruction directly as we need to shuffle
individual bits, but the smallest data we can operate on are bytes. We
therefore encrypt $8n$ messages at once which allows us to create bytewise
groupings. These messages are put into $4m$ registers with register $R_{4i}$
containing $S_0$, register $R_{4i+1}$ containing $S_1$ and so forth. With block
size $BS$ and register size $RS$, the following must hold:

\[
    8n\cdot BS=4m\cdot RS
\]

In the case of GIFT-64 with $BS=64$ and ARM NEON with $RS=128$, we get

\[
    8n\cdot 64=4m\cdot 128\Leftrightarrow n=m
\]

$n=m=1$ would be a valid choice which yields eight messages divided into four
registers. We choose $n=m=2$ so we can directly utilize the algorithm for bit
packing presented by the original GIFT authors, although it is simple to adapt
this algorithm to only four registers and eight messages by adjusting the
\texttt{SWAPMOVE} shift and mask values. Encrypting multiple blocks in parallel
is a realistic use case since usually a lot of blocks are encrypted under the
same key.

\subsubsection{Packing the data into bitslice format}

Let $a,b,\dots,p$ be sixteen messages of length $64$ with subscripts denoting
individual bits. We first put these messages into eight SIMD registers
$V_0,V_1,\dots,V_7$:

\begin{alignat*}{3}
    V_0&=b||a\qquad V_4&&=j||&&i \\
    V_1&=d||c\qquad V_5&&=l||&&k \\
    V_2&=f||e\qquad V_6&&=n||&&m \\
    V_3&=h||g\qquad V_7&&=p||&&o
\end{alignat*}

We then use the \texttt{SWAPMOVE} technique to bring the data into bitslice
format. This operation operates on two registers $A,B$ using mask $M$ and shift
value $N$. It swaps bits in $A$ masked by $(M\ll N)$ with bits in $B$ masked by
$M$ in using only three XOR-, one AND- and two shift operations.

\begin{align*}
    &\text{SWAPMOVE}(A,B,M,N): \\
    &\qquad T=((A\gg N)\oplus B)\land M \\
    &\qquad B=B\oplus T \\
    &\qquad A=A\oplus (T\ll N) \\
\end{align*}

% TODO change shr/shl implementation to only use vector registers
One caveat of this approach is the fact that NEON registers cannot be shifted
in their entirety due to the fact bits are not able to cross lanes. This leads
to the problem of being able to shift at most two lanes of 64 bits at once. We
thus need to implement the \texttt{shr(V,n)} and \texttt{shl(V,n)} operations
on our own. This can be done by extracting and shifting the overflow $ov$ out
of $V=V[1]||V[0]$, shifting the lanes individually and finally ORing the
overflow bits to the corresponding vector element.

\begin{alignat*}{2}
    &\text{shl}(V,n): \\
    &\qquad ov&&=V[0]\gg_{64-n} \\
    &\qquad V[0]&&=V[0]\ll \\
    &\qquad V[1]&&=(V[1]\ll)\lor ov
\end{alignat*}

The following operations group all $i$-th bits of the messages $a,c,\dots,o$
into bytes and put these into the lower half of the registers $V_{i\bmod 8}$.
The same is done for messages $b,d,\dots,p$, only differing in that the bytes
are put into the upper half of the registers.

\begin{align*}
    &\text{SWAPMOVE}(V_0,V_1,0x5555\dots 55,1) &\text{SWAPMOVE}(V_4,V_5,0x5555\dots 55,1) \\
    &\text{SWAPMOVE}(V_2,V_3,0x5555\dots 55,1) &\text{SWAPMOVE}(V_6,V_7,0x5555\dots 55,1) \\
    &\text{SWAPMOVE}(V_0,V_2,0x3333\dots 33,2) &\text{SWAPMOVE}(V_4,V_6,0x3333\dots 33,2) \\
    &\text{SWAPMOVE}(V_1,V_3,0x3333\dots 33,2) &\text{SWAPMOVE}(V_5,V_7,0x3333\dots 33,2) \\
    &\text{SWAPMOVE}(V_0,V_4,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_1,V_5,0x0f0f\dots 0f,4) \\
    &\text{SWAPMOVE}(V_2,V_6,0x0f0f\dots 0f,4) &\text{SWAPMOVE}(V_3,V_7,0x0f0f\dots 0f,4)
\end{align*}

With $Ax=o_xm_xk_xj_xg_xe_xc_xa_x$ and $Bx=p_xn_xl_xi_xh_xf_xd_xb_x$ denoting
byte groups, our data now has the following permutation-friendly format:

\scriptsize
\[
    \begin{array}{c|llllllll|llllllll}
        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
        \hline
        V_0 & B56 & B48 & B40 & B32 & B24 & B16 & B8  & B0 & A56 & A48 & A40 & A32 & A24 & A16 & A8  & A0 \\
        V_1 & B57 & B49 & B41 & B33 & B25 & B17 & B9  & B1 & A57 & A49 & A41 & A33 & A25 & A17 & A9  & A1 \\
        V_2 & B58 & B50 & B42 & B34 & B26 & B18 & B10 & B2 & A58 & A50 & A42 & A34 & A26 & A18 & A10 & A2 \\
        V_3 & B59 & B51 & B43 & B35 & B27 & B19 & B11 & B3 & A59 & A51 & A43 & A35 & A27 & A19 & A11 & A3 \\
        V_4 & B60 & B52 & B44 & B36 & B28 & B20 & B12 & B4 & A60 & A52 & A44 & A36 & A28 & A20 & A12 & A4 \\
        V_5 & B61 & B53 & B45 & B37 & B29 & B21 & B13 & B5 & A61 & A53 & A45 & A37 & A29 & A21 & A13 & A5 \\
        V_6 & B62 & B54 & B46 & B38 & B30 & B22 & B14 & B6 & A62 & A54 & A46 & A38 & A30 & A22 & A14 & A6 \\
        V_7 & B63 & B55 & B47 & B39 & B31 & B23 & B15 & B7 & A63 & A55 & A47 & A39 & A31 & A23 & A15 & A7
    \end{array}
\]
\normalsize

Although this would already work, we prefer to have only bits of the same
messages in each register - otherwise the permutation would need to operate on
two source registers with the added requirement of storing the pre-permutation
values for the first four registers, slowing down the round function through
superfluous load/stores. This transformation is trivial by use of
\texttt{TBL} with two data source operands. The final data format we operate on
is as follows:

\[
    \scriptsize
    \begin{array}{c|llllllll|llllllll}
        n & 15 & 14 & 13 & 12 & 11 & 10 & 9 & 8 & 7 & 6 & 5 & 4 & 3 & 2 & 1 & 0 \\
        \hline
        V_0 & A60 & A56 & A52 & A48 & A44 & A40 & A36 & A32 & A28 & A24 & A20 & A16 & A12 & A8  & A4 & A0 \\
        V_1 & A61 & A57 & A53 & A49 & A45 & A41 & A37 & A33 & A29 & A25 & A21 & A17 & A13 & A9  & A5 & A1 \\
        V_2 & A62 & A58 & A54 & A50 & A46 & A42 & A38 & A34 & A30 & A26 & A22 & A18 & A14 & A10 & A6 & A2 \\
        V_3 & A63 & A59 & A55 & A51 & A47 & A43 & A39 & A35 & A31 & A27 & A23 & A19 & A15 & A11 & A7 & A3 \\
        V_4 & B60 & B56 & B52 & B48 & B44 & B40 & B36 & B32 & B28 & B24 & B20 & B16 & B12 & B8  & B4 & B0 \\
        V_5 & B61 & B57 & B53 & B49 & B45 & B41 & B37 & B33 & B29 & B25 & B21 & B17 & B13 & B9  & B5 & B1 \\
        V_6 & B62 & B58 & B54 & B50 & B46 & B42 & B38 & B34 & B30 & B26 & B22 & B18 & B14 & B10 & B6 & B2 \\
        V_7 & B63 & B59 & B55 & B51 & B47 & B43 & B39 & B35 & B31 & B27 & B23 & B19 & B15 & B11 & B7 & B3
    \end{array}
\]

We can now create permutation tables using the specification of the individual
slice permutations $P_i$ which are then applied to $V_i$ and $V_{i+4}$
respectively:

\[
    \begin{array}{c|llllllllllllllll}
        j & 0 & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 & 10 & 11 & 12 & 13 & 14 & 15 \\
        \hline
        P_0(j) & 0 & 12 & 8 & 4 & 1 & 13 & 9 & 5 & 2 & 14 & 10 & 6 & 3 & 15 & 11 & 7 \\
        P_1(j) & 4 & 0 & 12 & 8 & 5 & 1 & 13 & 9 & 6 & 2 & 14 & 10 & 7 & 3 & 15 & 11 \\
        P_2(j) & 8 & 4 & 0 & 12 & 9 & 5 & 1 & 13 & 10 & 6 & 2 & 14 & 11 & 7 & 3 & 15 \\
        P_3(j) & 12 & 8 & 4 & 0 & 13 & 9 & 5 & 1 & 14 & 10 & 6 & 2 & 15 & 11 & 7 & 3
    \end{array}
\]

One thing to take note of is the original permutation values only show where a
given byte should land, not which byte belongs to a certain position - i.e. for
$P_0$, byte 1 should land in position 12, but the byte belonging to position 1
is byte 4. Because \texttt{TBL} works in the latter way, we have to do some
trivial rearrangements.

Assuming the correct permutation values are put into registers
$V_8,V_9,V_{10},V_{11}$, this now allows us to compute the permutation layer
for all 16 blocks in only eight permutation instructions.

\begin{alignat*}{2}
    &\texttt{TBL V0, V0, V8}\qquad &&\texttt{TBL V1, V1, V9} \\
    &\texttt{TBL V4, V4, V8}\qquad &&\texttt{TBL V5, V5, V9} \\
    &\texttt{TBL V2, V2, V10}\qquad &&\texttt{TBL V3, V3, V11} \\
    &\texttt{TBL V6, V6, V10}\qquad &&\texttt{TBL V7, V7, V11}
\end{alignat*}

\subsubsection{Round key function}

In contrast to packing and unpacking of data which is only done once in the
beginning and end, a round key is derived for every round, so the round key
derivation function needs to be as fast as possible. A simple but naive
approach for one round would be to generate a single round key, copy it 15
times and pack the resulting registers similar to how we proceed with the
messages. Due to the cost of packing the messages, this is prohibitivly
expensive. Because we know where each byte group ends up after packing, we can
directly XOR the round key bits to the correct position. Extending these bits
to bytes can then be done simply by repeatedly shifting and ORing the registers
together.

\section{Strategies for Camellia}

\subsection{Optimized non-SIMD}

The original paper proposes various platform-independent ways to implement
Camellia efficiently \cite{camellia:2001}. Only some of these apply to the
ARMv8 architecture since features like the inline barrel shifter and bitfield
manipulation instructions generally offer better performance.

\begin{description}
    \item[XOR cancellation property in key schedule:]
        While deriving $K_A$ in the key schedule, the instruction sequence

        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0), \Sigma_1) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

        causes cancellations, allowing us to eliminate some operations by
        replacing it with the following:

        \begin{align*}
            K_{A_L}&\leftarrow F(K_{L_R}\oplus F(K_{L_L}, \Sigma_0)) \\
            K_{A_R}&\leftarrow F(K_{L_L}\oplus \Sigma_1)
        \end{align*}

    \item[Absorption of whitening keys:] Whitening keys $kw_1,kw_3$ can be
        absorbed into other subkeys to save two XOR operations

    \item[Efficiently computing $(P\circ S)$:] This technique applies to 64-bit processors.
        By preparing tables

        \begin{alignat*}{9}
            & SP_0(y_{0(8)})=(&s_0(y_0),&& s_0(y_0),&& s_0(y_0),&& 0,&& s_0(y_0),&& 0,&& 0,&& s_0(y_0)&) \\
            & SP_1(y_{1(8)})=(&0,&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& s_1(y_1),&& 0,&& 0&) \\
            & SP_2(y_{2(8)})=(&s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0,&& s_2(y_2),&& s_2(y_2),&& 0&) \\
            & SP_3(y_{3(8)})=(&s_3(y_3),&& s_3(y_3),&& 0,&& s_3(y_3),&& 0,&& 0,&& s_3(y_3),&& s_3(y_3)&) \\
            & SP_4(y_{4(8)})=(&0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4),&& 0,&& s_2(y_4),&& s_2(y_4),&& s_2(y_4)&) \\
            & SP_5(y_{5(8)})=(&s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5),&& s_3(y_5),&& 0,&& s_3(y_5),&& s_3(y_5)&) \\
            & SP_6(y_{6(8)})=(&s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6),&& s_4(y_6),&& s_4(y_6),&& 0,&& s_4(y_6)&) \\
            & SP_7(y_{7(8)})=(&s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0,&& s_1(y_7),&& s_1(y_7),&& s_1(y_7),&& 0&)
        \end{alignat*},

        we can compute $(P\circ S)(X_{(64)})$ using only 8 table lookups and 7
        XORs in the following way:

        \[
            (z_1', z_2', z_3', z_4', z_5', z_6', z_7', z_8')\leftarrow\bigoplus_{i=0}^7 SP_i(y_i)
        \]

\end{description}

\subsection{Byteslicing}

Because Camellia is a byte-oriented block cipher, we can pursue a similar
strategy as for GIFT: find a convenient data packing format and a way to apply
necessary operations in parallel.

We choose a bytesliced representation by encrypting 16 plaintext blocks at once
so we can fill 16 vector registers with register $V_i$ containing all $i$-th
bytes. This lends itself well to a NEON implementation since S-boxes as well as
the FL layer, Feistel round and packing/unpacking functions can be implemented
efficiently.

\subsubsection{Packing and unpacking}

Packing and unpacking of data can be done efficiently by use of 32 \texttt{TBL}
instructions and is summarized in Figure \ref{fig:byteslicing}. After
packing, every $i$-th register will contain all $i$-th bytes of the 16 input
blocks. Unpacking is done in a similar way with different permutation tables.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{Figures/byteslicing.pdf}
    \caption{Camellia bytesliced packing function with 32 \texttt{TBL} operations.
    Curly brackets represent a 4-element vector array being used as source registers.}
    \label{fig:byteslicing}
\end{figure}

\subsubsection{Hardware-accelerated Camellia S-box}

A bytesliced implementation strategy for the S-box on ARMv8 can be derived
from already existing x86-optimized implementations utilizing the AES-NI
advanced encryption standard instruction set \cite{bcfastimplx86:2013}. NEON
itself possesses cryptographic extensions for finite field arithmetic and AES
as well as SHA calculations. These can be used to produce an accelerated
Camellia implementation due to the algebraic similarity of the AES- and
Camellia S-boxes.

The AES S-box works by multiplicatively inverting $x\in\F{8}$ over GF$(2^8)$ and then
applying an affine transformation $A$:

\[
    s(x):\F{8}\rightarrow\F{8}, x\mapsto A(x^{-1}_{\text{GF}(2^8)})
\]

The Camellia S-box $s_0$ is defined as follows:

\[
    s_0(x):\F{8}\rightarrow\F{8}, x\mapsto h(g(f(x\oplus 0xc5)))\oplus 0x6e
\]

with affine transformations $h,f$ and the multiplicative inversion function $g$
in the composite field GF$((2^4)^2)$. Because Galois fields of equal size are
isomorphic, there exist affine isomorphisms $\delta,\delta^{-1}$ between
GF($2^8$) and GF($(2^4)^2$) respectively \cite{satohmorioka:2003}:

\begin{alignat*}{2}
    \delta&:\text{GF}(2^8)&&\rightarrow\text{GF}((2^4)^2) \\
    \delta^{-1}&:\text{GF}((2^4)^2)&&\rightarrow\text{GF}(2^8)
\end{alignat*}

NEON provides the \texttt{AESE} instruction for one round of AES encryption
which works on a 128-bit vector register. By applying the inverse of ShiftRow
to x, we can apply the AES S-box to 16 bytes at once.

\[
    \texttt{AESE}(x)=\text{SubBytes}(\text{ShiftRow(x)})\Leftrightarrow \texttt{AESE}(\text{ShiftRow}^{-1}(x))=\text{SubBytes}(x)
\]

We can then reverse the affine transformation $A$ due to bijectivity and
extract the multiplicative inverse of all 16 vector bytes in GF$(2^8)$.

\[
    x^{-1}_{\text{GF}(2^8)}=A^{-1}(A(x^{-1}_{\text{GF}(2^8)}))=A^{-1}(\text{SubBytes}(x))=A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(x)))
\]

Using the affine isomorphism, we transform this inverse into the inverse in
GF$((2^4)^2)$ which is equal to $g(x)$:

\[
    g(x)=x^{-1}_{\text{GF}((2^4)^2)}=\delta(x^{-1}_{\text{GF}(2^8)})=\delta(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(x))))
\]

We now redefine $h,f$ as $H,F$ such that they include addition of constants
$0xc5$ and $0x6e$. We can now use the inverse in conjunction with $H$ and $F$
and an additional input transformation to calculate the Camellia S-box:

\begin{align*}
    s_0(x)&=h(g(f(x\oplus 0xc5)))\oplus 0x6e \\
          &=H(g(F(x))) \\
          &=H(\delta(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(\delta^{-1}(F(x)))))))
\end{align*}

Because of different bit endianness of the matrices representing
$\delta,\delta^{-1}$, we define an additional bit-swapping function:

\[
    S=S^{-1}:\begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}\mapsto
    \begin{pmatrix}
        0 & 0 & 0 & 0 & 0 & 0 & 0 & 1 \\
        0 & 0 & 0 & 0 & 0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 0 & 0 & 1 & 0 & 0 & 0 \\
        0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 \\
        1 & 0 & 0 & 0 & 0 & 0 & 0 & 0
    \end{pmatrix}
    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}
\]

The final equation serving as the basis for our implementation then becomes

\[
    s_0(x)=S^{-1}(H(\delta(S(A^{-1}(\texttt{AESE}(\text{ShiftRow}^{-1}(S^{-1}(\delta^{-1}(F(S(x))))))))))
\]

While appearing to be computationally too intensive for any performance gains,
we shall notice that all of the transformations are affine and can therefore be
combined into a single operation by simple matrix multiplication. We combine
them into a pre-filter function $\theta_0$ and a post-filter function $\psi_0$:

\begin{align*}
    \theta_0(x)&=S^{-1}(\delta^{-1}(F(S(x)))) \\
    \psi_0(x)&=S^{-1}(H(\delta(S(A^{-1}(x)))))
\end{align*}

This simplifies our final equation:

\[
    s_0(x)=\psi_0(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x))))
\]

The other S-boxes $s_1,s_2,s_3$ are defined in terms of $s_0$ and rotations. We
can include these rotations by constructing additional S-box-specific filters
$\theta_3,\psi_1,\psi_2$:

\begin{alignat*}{2}
    s_1(x)&=s_0(x)\lll_1&&=\psi_1(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x)))) \\
    s_2(x)&=s_0(x)\ggg_1&&=\psi_2(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_0(x)))) \\
    s_3(x)&=s_0(x\lll_1)&&=\psi_0(\texttt{AESE}(\text{ShiftRow}^{-1}(\theta_3(x))))
\end{alignat*}

\subsubsection{Fast matrix multiplication}

Efficient application of the filter functions is essential to performance and
can be implemented in parallel by use of the \texttt{TBL} instruction. First
note that matrix-vector multiplication $\F{8\times 8}\cdot\F{8}\rightarrow
\F{8}$ can be decomposed into two multiplications $\F{8\times
4}\cdot\F{4}\rightarrow \F{8}$ with subsequent addition of the results:

\[
    \begin{pmatrix}
        1 & 1 & 1 & 0 & 1 & 1 & 0 & 1 \\
        0 & 0 & 1 & 1 & 0 & 0 & 1 & 0 \\
        1 & 1 & 0 & 1 & 0 & 0 & 0 & 0 \\
        1 & 0 & 1 & 1 & 0 & 0 & 1 & 1 \\
        0 & 0 & 0 & 0 & 1 & 1 & 0 & 0 \\
        1 & 0 & 1 & 0 & 0 & 1 & 0 & 0 \\
        0 & 0 & 1 & 0 & 1 & 1 & 0 & 0 \\
        1 & 0 & 0 & 0 & 0 & 1 & 1 & 0 \\
    \end{pmatrix}\cdot
    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\\x_4\\x_5\\x_6\\x_7\end{pmatrix}=
    \begin{pmatrix}
        1 & 1 & 0 & 1 \\
        0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 0 \\
        0 & 0 & 1 & 1 \\
        1 & 1 & 0 & 0 \\
        0 & 1 & 0 & 0 \\
        1 & 1 & 0 & 0 \\
        0 & 1 & 1 & 0 \\
    \end{pmatrix}\cdot
    \begin{pmatrix}x_0\\x_1\\x_2\\x_3\end{pmatrix}\oplus
    \begin{pmatrix}
        1 & 1 & 1 & 0 \\
        0 & 0 & 1 & 1 \\
        1 & 1 & 0 & 1 \\
        1 & 0 & 1 & 1 \\
        0 & 0 & 0 & 0 \\
        1 & 0 & 1 & 0 \\
        0 & 0 & 1 & 0 \\
        1 & 0 & 0 & 0 \\
    \end{pmatrix}\cdot
    \begin{pmatrix}x_4\\x_5\\x_6\\x_7\end{pmatrix}
\]

We can then tabulate the decomposed multiplications since each input now
consists of only four bits which allows us to fit the 16 possible 8-bit results
into a single vector register. A single matrix multiplication is then executed
for 16 bytes in parallel by use of two \texttt{TBL} and one XOR instruction with
an additional AND and SHR operation for masking off the lower/upper 4 bits.

\subsubsection{Permutation}

Since all bytes of the same position are collected in the same register,
applying the permutation can be realized in 16 XOR operations as described in
the Camellia specification with the difference that we choose to not swap the
lower and higher 4 bytes, but rather compensate for this by choosing the
correct bytes when XORing with the other half later in the Feistel round.

\begin{figure}[h!]
    \centering
    \includegraphics[width=\textwidth]{Figures/camelliaround.pdf}
    \caption{Camellia bytesliced Feistel round function}
\end{figure}

\subsubsection{FL layer}

The two functions $FL,FL^{-1}$ can be implemented in a straightforward manner
once 1-bit left rotation has been defined for the bytesliced representation.
This can be facilitated by extracting the highest bit of each byte, shifting it
all the way to the right and storing it as overflow $ov_i$. $ov_i$ is then added
to the left-shifted value of byte $(i + 4) (\bmod 4)$ for the final result.

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.8\textwidth]{Figures/camellia_bytesliced_rol.pdf}
    \caption{Camellia bytesliced 1-bit left rotate}
\end{figure}